---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
################################################################################
# DEFAULT SETTINGS
################################################################################
knitr::opts_chunk$set(message = FALSE) # don't knit messages
knitr::opts_chunk$set(warning = FALSE) # don't knit warnings
```

# clusterTruster

This is a package to benchmark linkages in hierarchical clustering. Please see our paper on arXiv for more detailed examples: https://arxiv.org/abs/1906.03336. The authors of the paper and this package are Xiao Hui Tai (https://xhtai.github.io/) and Kayla Frisoli (http://stat.cmu.edu/~kfrisoli/).

## Installation

You can install `clusterTruster` from github with:

```{r gh-installation, eval = FALSE}
devtools::install_github("xhtai/clusterTruster")
library(clusterTruster)
```

## Examples

### iris

We can analyze the iris data set as follows. We first center and scale the four features, "Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", and generate the l2 distance between all pairs. We then use various linkage methods to generate all associated evaluation metrics.

```{r, eval = TRUE}
myPairwise <- clusterTruster::genPairwise(iris, 5)
iris[, 1:4] <- scale(iris[, 1:4])
myPairwise <-
  clusterTruster::genSimDiff(iris, 1:4, myPairwise, 1:2, "l2dist")
```


```{r, eval = TRUE, cache=TRUE}
outMetrics_iris_single <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "single"
)
outMetrics_iris_complete <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "complete"
)
outMetrics_iris_average <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "average"
)
outMetrics_iris_centroid <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "centroid"
)
outMetrics_iris_minimax <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "minimax"
)
```

#### Plotting linkage method results for iris

To compare single, complete, and average:

```{r, fig.height=3.5, fig.width=10, fig.align='center'}
clusterTruster::plotResultsGG_base(
  single = outMetrics_iris_single,
  complete = outMetrics_iris_complete,
  average = outMetrics_iris_average,
  correlation = FALSE,
  write_plot = FALSE,
  plot_type = c("all", "minimax", "misclass", "pr")
)
```

To compare all:

```{r, fig.height=3.5, fig.width=10, fig.align='center'}
clusterTruster::plotResultsGG_base(
  single = outMetrics_iris_single,
  complete = outMetrics_iris_complete,
  average = outMetrics_iris_average,
  centroid = outMetrics_iris_centroid,
  minimax = outMetrics_iris_minimax,
  correlation = FALSE,
  write_plot = FALSE,
  plot_type = c("all", "minimax", "misclass", "pr")
)
```


# Reproducing Paper Results 

### NBIDE
The package also contains an example data set, called `removedDups`. This analyzes 144 images of cartridge cases, taken from NIST's [Ballistics Toolmarks Research Database](https://tsapps.nist.gov/NRBTD). There are 12 images each taken from 12 different guns, from the NBIDE study. Pairwise comparisons have been pre-computed using the [`cartridges3D` package](https://github.com/xhtai/cartridges3D), and stored in `removedDups`. In this case, we do not need to use the functions `genPairwise()` or `genSimDiff()`, and can simply input the appropriate data and column information into `getMetrics`, as follows.

```{r, eval = FALSE}
NBIDEminimax <- getMetrics(
  removedDups,
  pairColNums = 1:2,
  matchColNum = "match",
  distSimCol = "corrMax",
  linkage = "minimax",
  myDist = FALSE
) 
```


## Creating outMetrics/ folder

## Plots

This is some example code for generating plots, assuming that metrics have been generated using each linkage method and stored in a Rdata files with their corresponding names. I have a folder called `outMetrics/` with the following contents: `faces_average.Rdata`, `faces_centroid.Rdata`, `faces_complete.Rdata`, `faces_minimax.Rdata`, `faces_single.Rdata`

```{r, eval = FALSE, echo=FALSE}
# non-Gggplot

# plot(single$numClusters, single$maxMinimax,
#      xlab = "Number of clusters", ylab = "Max minimax radius",
#      main = "Max minimax radius",
#      type = "l")
# lines(complete$numClusters, complete$maxMinimax, col = 2)
# lines(average$numClusters, average$maxMinimax, col = 3)
# lines(centroid$numClusters, centroid$maxMinimax, col = 4)
# lines(minimax$numClusters, minimax$maxMinimax, col = 5)
# legend("topright", legend = c("Single", "Complete", "Average", "Centroid", "Minimax"), col = 1:5, lty = c(1, 1, 1, 1, 1))
```

```{r, eval = FALSE}
plotResultsGG("faces", plot_type = "minimax")
plotResultsGG("faces", plot_type = "misclass")
plotResultsGG("faces", plot_type = "pr")
plotResultsGG("faces", write_plot = FALSE, plot_type = "all")
```

For the following, I would need to have `outMetrics/` with the following contents: `NBIDE_average.Rdata`, `NBIDE_centroid.Rdata`, `NBIDE_complete.Rdata`, `NBIDE_minimax.Rdata`, `NBIDE_single.Rdata`

```{r, eval = FALSE}
plotResultsGG("NBIDE", correlation = TRUE, write_plot = FALSE, plot_type = "all")
```


## License
The `clusterTruster` package is licensed under GPLv3 (http://www.gnu.org/licenses/gpl.html).
