---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
################################################################################
# DEFAULT SETTINGS
################################################################################
knitr::opts_chunk$set(message = FALSE) # don't knit messages
knitr::opts_chunk$set(warning = FALSE) # don't knit warnings
```

# clusterTruster

This is a package to benchmark linkages in hierarchical clustering. Please see our paper on arXiv for more detailed examples: https://arxiv.org/abs/1906.03336. The authors of the paper and this package are Xiao Hui Tai (https://xhtai.github.io/) and Kayla Frisoli (http://stat.cmu.edu/~kfrisoli/).

## Installation

You can install `clusterTruster` from github with:

```{r gh-installation, eval = FALSE}
devtools::install_github("xhtai/clusterTruster")
library(clusterTruster)
```

## Examples

### iris

We can analyze the iris data set as follows. We first center and scale the four features, "Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", and generate the l2 distance between all pairs. We then use various linkage methods to generate all associated evaluation metrics.

```{r, eval = TRUE}
myPairwise <- clusterTruster::genPairwise(iris, 5)
iris[, 1:4] <- scale(iris[, 1:4])
myPairwise <-
  clusterTruster::genSimDiff(iris, 1:4, myPairwise, 1:2, "l2dist")
```


```{r, eval = TRUE, cache=TRUE}
outMetrics_iris_single <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "single"
)
outMetrics_iris_complete <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "complete"
)
outMetrics_iris_average <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "average"
)
outMetrics_iris_centroid <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "centroid"
)
outMetrics_iris_minimax <- clusterTruster::getMetrics(
  myPairwise,
  pairColNums = 1:2,
  matchColNum = 3,
  distSimCol = "l2dist",
  linkage = "minimax"
)
```

#### Plotting linkage method results for iris

To compare single, complete, and average:

```{r, fig.height=3.5, fig.width=10, fig.align='center'}
clusterTruster::plotResultsGG_base(
  single = outMetrics_iris_single,
  complete = outMetrics_iris_complete,
  average = outMetrics_iris_average,
  correlation = FALSE,
  write_plot = FALSE,
  plot_type = c("all", "minimax", "misclass", "pr")
)
```

To compare all:

```{r, fig.height=3.5, fig.width=10, fig.align='center'}
clusterTruster::plotResultsGG_base(
  single = outMetrics_iris_single,
  complete = outMetrics_iris_complete,
  average = outMetrics_iris_average,
  centroid = outMetrics_iris_centroid,
  minimax = outMetrics_iris_minimax,
  correlation = FALSE,
  write_plot = FALSE,
  plot_type = c("all", "minimax", "misclass", "pr")
)
```


## Reproducing Paper Results 
To reproduce the results in our paper, Benchmarking Minimax Linkage in Hierarchical Clustering (https://arxiv.org/abs/1906.03336), refer to this vignette.

## License
The `clusterTruster` package is licensed under GPLv3 (http://www.gnu.org/licenses/gpl.html).
