---
title: "Reproducing Paper Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reproducePaperResults}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(clusterTruster)
```

This vignette demonstrates how to reproduce the results in our paper, Benchmarking Minimax Linkage in Hierarchical Clustering (https://arxiv.org/abs/1906.03336), using the `clusterTruster` package. The following description is brief; for full details refer to the paper.

## Data sets
For the benchmark study, we use a total of 7 data sets. For the observations in each data set, we need to define a distance metric in order to run hierarchical clustering. The data sets and distance metrics used are shown below.

| Data set | Distance metric |
|:---------|:------|
|  Olivetti faces  |   $\ell_2$ distance |
|  Colon cancer |   Correlation  |
|    Prostate cancer |   Correlation  |
|    Iris |   $\ell_2$ distance  |
|    NBIDE |   Maximum cross-correlation  |
|    FBISW |   Maximum cross-correlation  |
|    Simulations: |     |
|    - Spherical |   $\ell_1, \ell_2$ distance   |
|    - Elliptical |   $\ell_1, \ell_2$ distance   |
|    - Outliers |   $\ell_1, \ell_2$ distance   |

## Perform clustering and generate comparisons between different linkage methods

The following is code used for each of the above data sets. The code shows the source of the data as well as how to generate comparison metrics.

### Olivetti faces
```{r, eval = FALSE}
library(RnavGraphImageData)
data("faces")

# https://rdrr.io/github/jlmelville/snedata/src/R/olivetti.R
olivetti_faces <- function() {
  if (!requireNamespace("RnavGraphImageData", quietly = TRUE,
                        warn.conflicts = FALSE)) {
    stop("olivetti_faces function requires 'RnavGraphImageData' package")
  }
  faces <- NULL
  utils::data("faces", envir = environment())
  
  df <- as.data.frame(t(faces))
  npeople <- 40
  nposes <- 10
  colnames(df) <- sapply(seq(1, 4096), function(x) { paste0("px", x)})
  rownames(df) <- apply(expand.grid(seq(1, nposes), seq(1, npeople)), 1,
                        function(x) { paste(x[2], x[1], sep = "_") })
  df$Label <-  factor(as.numeric(cut(1:nrow(df), npeople)))
  
  df
}
faces <- olivetti_faces()
myPairwise <- genPairwise(datasetName = faces, labelColNum = 4097)
myPairwise <- genSimDiff(datasetName = faces, featureCols = 1:4096, allPairwise = myPairwise, pairColNums = 1:2, measure = "l2dist") 

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l2dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/faces", "_", linkages[l], ".Rdata"))
}
```


### Colon cancer
```{r, eval = FALSE}
library(HiDimDA)
data(AlonDS)

myPairwise <- genPairwise(datasetName = AlonDS, labelColNum = 1)
myPairwise <- genSimDiff(datasetName = AlonDS, featureCols = 2:2001, myPairwise, pairColNums = 1:2, "correlation")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "correlation", myDist = FALSE)
  save(outMetrics, file = paste0("./outMetrics/AlonDS", "_", linkages[l], ".Rdata"))
}
```

### Prostate cancer
There are various possible sources of these data. The version that we use can be accessed from within the package using `data(singh2002)`. It is downloaded from https://stat.ethz.ch/~dettling/bagboost.html. Do not use the data from the `sda` package in R, as this is scaled differently.

```{r, eval = FALSE}
data(singh2002) # 102x6034

myPairwise <- genPairwise(datasetName = singh2002, labelColNum = 6034)
myPairwise <- genSimDiff(datasetName = singh2002, featureCols = 1:6033, myPairwise, pairColNums = 1:2, "correlation")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "correlation", myDist = FALSE)
  save(outMetrics, file = paste0("./outMetrics/singh2002", "_", linkages[l], ".Rdata"))
}
```

### Iris
This is the famous `iris` data set that is included within R.

```{r, eval = FALSE}
data(iris)
myPairwise <- genPairwise(datasetName = iris, labelColNum = 5)
iris[, 1:4] <- scale(iris[, 1:4])
myPairwise <- genSimDiff(datasetName = iris, featureCols = 1:4, allPairwise = myPairwise, pairColNums = 1:2, measure = "l2dist")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l2dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/iris", "_", linkages[l], ".Rdata"))
}
```


### NBIDE
This data set is of 144 images of cartridge cases, taken from NIST's [Ballistics Toolmarks Research Database](https://tsapps.nist.gov/NRBTD). There are 12 images each taken from 12 different guns, from the NBIDE study. Pairwise comparisons have been pre-computed using the [`cartridges3D` package](https://github.com/xhtai/cartridges3D), and stored in `removedDups`.

In this case, we do not need to use the functions `genPairwise()` or `genSimDiff()`, and can simply input the appropriate data and column information into `getMetrics`, as follows.

```{r, eval = FALSE}
linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = removedDups, pairColNums = 1:2, matchColNum = "match", distSimCol = "corrMax", linkage = linkages[l], myDist = FALSE)
  save(outMetrics, file = paste0("./outMetrics/NBIDE", "_", linkages[l], ".Rdata"))
}
```

### FBISW
Similarly to NBIDE, data from FBISW have been precomputed, and are stored in `FBISW`.

```{r, eval = FALSE}
linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = FBISW, pairColNums = 1:2, matchColNum = "match", distSimCol = "corrMax", linkage = linkages[l], myDist = FALSE)
  save(outMetrics, file = paste0("./outMetrics/FBISW", "_", linkages[l], ".Rdata"))
}
```

### Simulations
```{r, eval = FALSE}
### 1. spherical
set.seed(7)
c1 <- MASS::mvrnorm(n = 100, rep(0, 10), Sigma = diag(10))
c2 <- MASS::mvrnorm(n = 100, c(2, 2, rep(0, 8)), Sigma = diag(10))
c3 <- MASS::mvrnorm(n = 100, c(0, 2, 2, rep(0, 7)), Sigma = diag(10))
spherical <- data.frame(clust = rep(1:3, each = 100), rbind(c1, c2, c3))

myPairwise <- genPairwise(datasetName = spherical, labelColNum = 1)
myPairwise <- genSimDiff(datasetName = spherical, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l2dist")
myPairwise <- genSimDiff(datasetName = spherical, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l1dist")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l2dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/sphericall2", "_", linkages[l], ".Rdata"))
  
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l1dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/sphericall1", "_", linkages[l], ".Rdata"))
}

### 2. elliptical
set.seed(7)
c1 <- MASS::mvrnorm(n = 100, rep(0, 10), Sigma = diag(x = c(1, 1, 1, 2, 2, 1, 1, 1, 1, 1)))
c2 <- MASS::mvrnorm(n = 100, c(2, 2, rep(0, 8)), Sigma = diag(x = c(1, 1, 1, 2, 2, 1, 1, 1, 1, 1)))
c3 <- MASS::mvrnorm(n = 100, c(0, 2, 2, rep(0, 7)), Sigma = diag(x = c(1, 1, 1, 2, 2, 1, 1, 1, 1, 1)))
elliptical <- data.frame(clust = rep(1:3, each = 100), rbind(c1, c2, c3))

myPairwise <- genPairwise(datasetName = elliptical, labelColNum = 1)
myPairwise <- genSimDiff(datasetName = elliptical, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l2dist")
myPairwise <- genSimDiff(datasetName = elliptical, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l1dist")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l2dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/ellipticall2", "_", linkages[l], ".Rdata"))
  
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l1dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/ellipticall1", "_", linkages[l], ".Rdata"))
}


### 3. outliers
set.seed(7)
c1 <- MASS::mvrnorm(n = 100, rep(0, 10), Sigma = diag(10))
c2 <- MASS::mvrnorm(n = 98, c(2, 2, rep(0, 8)), Sigma = diag(10))
c2out <- MASS::mvrnorm(n = 2, c(5, 5, rep(0, 8)), Sigma = diag(10))
c3 <- MASS::mvrnorm(n = 98, c(0, 2, 2, rep(0, 7)), Sigma = diag(10))
c3out <- MASS::mvrnorm(n = 2, c(0, 5, 5, rep(0, 7)), Sigma = diag(10))
outliers <- data.frame(clust = rep(1:3, each = 100), rbind(c1, c2, c2out, c3, c3out))

myPairwise <- genPairwise(datasetName = outliers, labelColNum = 1)
myPairwise <- genSimDiff(datasetName = outliers, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l2dist")
myPairwise <- genSimDiff(datasetName = outliers, featureCols = 2:11, allPairwise = myPairwise, pairColNums = 1:2, measure = "l1dist")

linkages <- c("single", "complete", "average", "centroid", "minimax")
for (l in 1:length(linkages)) {
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l2dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/outliersl2", "_", linkages[l], ".Rdata"))
  
  outMetrics <- getMetrics(allPairwise = myPairwise, pairColNums = 1:2, matchColNum = 3, linkage = linkages[l], distSimCol = "l1dist", myDist = TRUE)
  save(outMetrics, file = paste0("./outMetrics/outliersl1", "_", linkages[l], ".Rdata"))
}
```



## Plots
This is some example code for generating plots, assuming that metrics have been generated using each linkage method and stored in a Rdata files with their corresponding names, as demonstrated above. Assuming we have a folder called `outMetrics/` with the following contents: `faces_average.Rdata`, `faces_centroid.Rdata`, `faces_complete.Rdata`, `faces_minimax.Rdata`, `faces_single.Rdata`, we can run the following code.


```{r, eval = FALSE}
plotResultsGG("faces", plot_type = "minimax")
plotResultsGG("faces", plot_type = "misclass")
plotResultsGG("faces", plot_type = "pr")
plotResultsGG("faces", write_plot = FALSE, plot_type = "all")
```

For the NBIDE data set, we can run the following code, if we have `outMetrics/` with the following contents: `NBIDE_average.Rdata`, `NBIDE_centroid.Rdata`, `NBIDE_complete.Rdata`, `NBIDE_minimax.Rdata`, `NBIDE_single.Rdata`.

```{r, eval = FALSE}
plotResultsGG("NBIDE", correlation = TRUE, write_plot = FALSE, plot_type = "all")
```
